---
title: "Physical functioning and physical fitness in glioma patients: SF36 part 3"
output: html_document
date: '2024-03-04'
output: html_document
---

#### Email: m.e.c.blom@amsterdamumc.nl
#### Purpose of script: 
Code analysis 

#### Reviewed by Maxine Gorter

#### Notes 
# Part 3 #
# 3.1 SF36 diagnosis; correlations (Pearson continue/independent sample t-test dichotomy/ANOVA categorical) 
# 3.2 SF36 stable; correlations (Pearson continue/independent sample t-test dichotomy/ANOVA categorical)

# 3.3 CIS20 diagnosis; correlaties
# 3.4 CIS20 stable; correlaties

# 3.5 SF36 difference diagnosis/stable; correlations
# 3.6 CIS20 verschil diagnose/stabiel; correlaties 



# Load libraries
```{r warning = FALSE, message = FALSE}
library(plyr);
library(tidyverse);
library(readr);
library(readxl);
library(haven);
library(stats);
library(dplyr);
library(dbplyr);
library(MatchIt);
library(ggplot2);
library(tidyr);
library(car);
library(dunn.test);
library(vcd);
library(lme4);
library(coin);
library(afex); 
library(nlme);
library(QuantPsyc);
```




###### Part 3.1 ######

##### 3.1 SF36 diagnosis; correlations (Pearson continue/independent sample t-test dichotomy/ANOVA categorical) 

## Select data 
```{r}
# Select SF36_PF diagnosis
df_diagnosis_SF36 <- df %>%
  filter(Time_moment == 1)
```

## Univariate analyses; check each variable separately 
# Age
```{r}
# Remove NA
df_clean_age <- na.omit(df_diagnosis_SF36[, c("age", "SF36_PF")])
# Calculate Spearman-correlation between SF36_PF and age, ignore NA
correlatie_SF36_age <- cor.test(df_clean_age$SF36_PF, df_clean_age$age, method = "spearman", use = "complete.obs")
# Print results
print(paste("Spearman-correlatie between SF36_PF and age:", correlatie_SF36_age))

rho <- correlatie_SF36_age$estimate
p_value <- correlatie_SF36_age$p.value

# Print results
print(paste("Spearman-correlatie tussen SF36_PF en leeftijd:", rho))
print(paste("P-waarde:", p_value))

```

# Sex
```{r}
# Convert to factor
df_diagnosis_SF36$sex <- factor(df_diagnosis_SF36$sex)

# T-test (non-parametric)
ttest_sex <- wilcox.test(SF36_PF ~ sex, data = df_diagnosis_SF36)
print(ttest_sex)

```

# Educational level
```{r}
# Convert to factor
df_diagnosis_SF36$education_old <- factor(df_diagnosis_SF36$education_old)

# Create new column with grouped values 
df_diagnosis_SF36 <- df_diagnosis_SF36 %>%
  mutate(education_grouped = case_when(
    education_old %in% c(1, 2) ~ "1-2",
    education_old %in% c(3, 4, 5) ~ "3-4-5",
    education_old %in% c(6, 7, 8) ~ "6-7-8"
  ))

# Convert to factor
df_diagnosis_SF36$education_grouped <- factor(df_diagnosis_SF36$education_grouped)

# ANOVA (non parametric)
anova_education <- kruskal.test(SF36_PF ~ education_grouped, data = df_diagnosis_SF36)
print(anova_education)
summary(anova_education)
table(df_diagnosis_SF36$education_grouped)
# Post hoc
Posthoc_education <- dunn.test(df_diagnosis_SF36$SF36_PF, df_diagnosis_SF36$education_grouped, method = "bonferroni")

```

# Histology
```{r}
# Convert to factor
df_diagnosis_SF36$histology <- factor(df_diagnosis_SF36$histology)

# ANOVA (non parametric)
anova_histology <- kruskal.test(SF36_PF ~ histology, data = df_diagnosis_SF36)
print(anova_histology)
summary(anova_histology)
table(df_diagnosis_SF36$histology)
# Post hoc
Posthoc_histology <- dunn.test(df_diagnosis_SF36$SF36_PF, df_diagnosis_SF36$histology, method = "bonferroni")

```

# Tumor location
```{r}
# Convert to factor
df_diagnosis_SF36$location <- factor(df_diagnosis_SF36$location)

# Create new column with grouped values 
df_diagnosis_SF36 <- df_diagnosis_SF36 %>%
  mutate(location_grouped = case_when(
    location %in% c(2) ~ "frontal",
    location %in% c(1, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 20, 21, 22, 23, 24, 25, 26) ~ "non-frontal"
  ))

# T-test (non-parametric)
ttest_location <- wilcox.test(SF36_PF ~ location_grouped, data = df_diagnosis_SF36)
print(ttest_location)

```

# Tumor lateralization (side)
```{r}
# Convert to factor
df_diagnosis_SF36$side <- factor(df_diagnosis_SF36$side)

# ANOVA
anova_side <- kruskal.test(SF36_PF ~ side, data = df_diagnosis_SF36)
print(anova_side)
summary(anova_side)
table(df_diagnosis_SF36$side)
# Post hoc
Posthoc_side <- dunn.test(df_diagnosis_SF36$SF36_PF, df_diagnosis_SF36$side, method = "bonferroni")

```

# WHO grade
```{r}
# Convert to factor
df_diagnosis_SF36$graad <- factor(df_diagnosis_SF36$graad)

# ANOVA
anova_grade <- kruskal.test(SF36_PF ~ graad, data = df_diagnosis_SF36)
print(anova_grade)
summary(anova_grade)
table(df_diagnosis_SF36$graad)
# Post hoc
Posthoc_grade <- dunn.test(df_diagnosis_SF36$SF36_PF, df_diagnosis_SF36$graad, method = "bonferroni")

```

# Type of surgery
```{r}
# T-test (non parametric)
ttest_surgery <- wilcox.test(SF36_PF ~ OK_, data = df_diagnosis_SF36)
print(ttest_surgery)

```

# Presence of epilepsy
```{r}
# T-test (non parametric)
ttest_epilepsy <- wilcox.test(SF36_PF ~ epilepsy_dich, data = df_diagnosis_SF36)
print(ttest_epilepsy)

```

# Use of dexamethasone
```{r}
# Convert to factor
df_diagnosis_SF36$dexa <- factor(df_diagnosis_SF36$dexa)

# T-test (non parametric)
ttest_dexa <- wilcox.test(SF36_PF ~ dexa, data = df_diagnosis_SF36)
print(ttest_dexa)

```

# KPS
```{r}
# KPS groups: = or < 70 and 80-90-100
df_diagnosis_SF36 <- df_diagnosis_SF36 %>%
  mutate(KPS_grouped = case_when(
    kps_total %in% c(80, 90, 100) ~ "80-90-100",
    kps_total %in% c(10, 20, 30, 40, 50, 60, 70) ~ "70 or lower"
  ))

# Convert to factor
df_diagnosis_SF36$KPS_grouped <- factor(df_diagnosis_SF36$KPS_grouped)

# T-test (non parametric)
ttest_KPS <- wilcox.test(SF36_PF ~ KPS_grouped, data = df_diagnosis_SF36)
print(ttest_KPS)

```

# NIHSS score
```{r}
# NIHSS groups: = 0 or 1-6 
df_diagnosis_SF36 <- df_diagnosis_SF36 %>%
  mutate(NIHSS_grouped = case_when(
    NIHSS_totalscore %in% c(0) ~ "no disabilities",
    NIHSS_totalscore %in% c(1,2,3,4,5,6,7) ~ "one or more disabilities"
  ))

# T-test (non-parametric)
ttest_NIHSS <- wilcox.test(SF36_PF ~ NIHSS_grouped, data = df_diagnosis_SF36)
print(ttest_NIHSS)

```

## Multivariate analysis 
```{r}
set.seed(123)

# Multiple regression with VIF (Variance Inflation Factor)
# Regression model
model_diagnosis <- lm(SF36_PF ~ age + sex + education_grouped + side + graad + KPS_grouped + dexa, data = df_diagnosis_SF36)
summary(model_diagnosis)

# VIF
vif_values_diagnosis <- vif(model_diagnosis)
print(vif_values_diagnosis)

# Collineariteit tussen twee histology & grade
v_cramer <- assocstats(table(df_diagnosis_SF36$histology, df_diagnosis_SF36$graad))$cramer
table(v_cramer)

```










###### Part 3.2 ######

##### 3.2 SF36 stable; correlations (Pearson continue/independent sample t-test dichotomy/ANOVA categorical)

## Select data
```{r}
# Select SF36_PF stable phase 
df_stable_SF36 <- df %>%
  filter(Time_moment == 2)
```

## Univariate analyses; check each variable separately 
# Age
```{r}
# Remove NA
df_clean_age <- na.omit(df_stable_SF36[, c("age", "SF36_PF")])
# Calculate Spearman-correlation between SF36_PF and age, ignore NA
correlatie_SF36_age <- cor.test(df_clean_age$SF36_PF, df_clean_age$age, method = "spearman", use = "complete.obs")
# Print results
print(paste("Spearman-correlatie between SF36_PF and age:", correlatie_SF36_age))

rho <- correlatie_SF36_age$estimate
p_value <- correlatie_SF36_age$p.value

# Print results
print(paste("Spearman-correlatie tussen SF36_PF en leeftijd:", rho))
print(paste("P-waarde:", p_value))

```

# Sex
```{r}
# Convert to factor
df_stable_SF36$sex <- factor(df_stable_SF36$sex)

# T-test (non-parametric)
ttest_sex <- wilcox.test(SF36_PF ~ sex, data = df_stable_SF36)
print(ttest_sex)

```

# Educational level
```{r}
# Convert to factor
df_stable_SF36$education_old <- factor(df_stable_SF36$education_old)

# Create new column with grouped values 
df_stable_SF36 <- df_stable_SF36 %>%
  mutate(education_grouped = case_when(
    education_old %in% c(1, 2) ~ "1-2",
    education_old %in% c(3, 4, 5) ~ "3-4-5",
    education_old %in% c(6, 7, 8) ~ "6-7-8"
  ))

# Convert to factor
df_stable_SF36$education_grouped <- factor(df_stable_SF36$education_grouped)

# ANOVA (non parametric)
anova_education <- kruskal.test(SF36_PF ~ education_grouped, data = df_stable_SF36)
print(anova_education)
summary(anova_education)
table(df_stable_SF36$education_grouped)
# Post hoc
Posthoc_education <- dunn.test(df_stable_SF36$SF36_PF, df_stable_SF36$education_grouped, method = "bonferroni")

```

# Histology
```{r}
# Convert to factor
df_stable_SF36$histology <- factor(df_stable_SF36$histology)

# ANOVA (non parametric)
anova_location <- kruskal.test(SF36_PF ~ histology, data = df_stable_SF36)
print(anova_location)
summary(anova_location)
table(df_stable_SF36$histology)
# Post hoc
Posthoc_location <- dunn.test(df_stable_SF36$SF36_PF, df_stable_SF36$histology, method = "bonferroni")

```

# Tumor location
```{r}
# Convert to factor
df_stable_SF36$location <- factor(df_stable_SF36$location)

# Create new column with grouped values 
df_stable_SF36 <- df_stable_SF36 %>%
  mutate(location_grouped = case_when(
    location %in% c(2) ~ "frontal",
    location %in% c(1, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 20, 21, 22, 23, 24, 25, 26) ~ "non frontal"
  ))

# T-test (non-parametric)
ttest_location <- wilcox.test(SF36_PF ~ location_grouped, data = df_stable_SF36)
print(ttest_location)

```

# Tumor lateralization (side)
```{r}
# Convert to factor
df_stable_SF36$side <- factor(df_stable_SF36$side)

# ANOVA
anova_side <- kruskal.test(SF36_PF ~ side, data = df_stable_SF36)
print(anova_side)
summary(anova_side)
table(df_stable_SF36$side)
# Post hoc
Posthoc_side <- dunn.test(df_stable_SF36$SF36_PF, df_stable_SF36$side, method = "bonferroni")

```

# WHO grade
```{r}
# Convert to factor
df_stable_SF36$graad <- factor(df_stable_SF36$graad)

# ANOVA
anova_grade <- kruskal.test(SF36_PF ~ graad, data = df_stable_SF36)
print(anova_grade)
summary(anova_grade)
table(df_stable_SF36$graad)
# Post hoc
Posthoc_grade <- dunn.test(df_stable_SF36$SF36_PF, df_stable_SF36$graad, method = "bonferroni")

```

# Type of surgery
```{r}
# T-test (non parametric)
ttest_surgery <- wilcox.test(SF36_PF ~ OK_, data = df_stable_SF36)
print(ttest_surgery)

```

# Presence of epilepsy
```{r}
# T-test (non parametric)
ttest_epilepsy <- wilcox.test(SF36_PF ~ epilepsy_dich, data = df_stable_SF36)
print(ttest_epilepsy)

```

# Use of dexamethasone
```{r}
# Convert to factor
df_stable_SF36$dexa <- factor(df_stable_SF36$dexa)

# T-test (non parametric)
ttest_dexa <- wilcox.test(SF36_PF ~ dexa, data = df_stable_SF36)
print(ttest_dexa)

```

# KPS
```{r}
# KPS groups: = or < 70 and 80-90-100
df_stable_SF36 <- df_stable_SF36 %>%
  mutate(KPS_grouped = case_when(
    kps_total %in% c(80, 90, 100) ~ "80-90-100",
    kps_total %in% c(10, 20, 30, 40, 50, 60, 70) ~ "70 or lower"
  ))

# Convert to factor
df_stable_SF36$KPS_grouped <- factor(df_stable_SF36$KPS_grouped)

# T-test (non parametric)
ttest_KPS <- wilcox.test(SF36_PF ~ KPS_grouped, data = df_stable_SF36)
print(ttest_KPS)

```

# Radiotherapy & chemotherapy
```{r}
# Convert to factor
df_stable_SF36$RT_XT_treatment <- factor(df_stable_SF36$RT_XT_treatment)

# ANOVA
anova_RT_XT_treatment  <- kruskal.test(SF36_PF ~ RT_XT_treatment , data = df_stable_SF36)
print(anova_RT_XT_treatment )
summary(anova_RT_XT_treatment )
table(df_stable_SF36$RT_XT_treatment )
# Post hoc
Posthoc_RT_XT_treatment  <- dunn.test(df_stable_SF36$SF36_PF, df_stable_SF36$RT_XT_treatment , method = "bonferroni")

```

# NIHSS score
```{r}
# NIHSS groups: = 0 or 1-6 
df_stable_SF36 <- df_stable_SF36 %>%
  mutate(NIHSS_grouped = case_when(
    NIHSS_totalscore %in% c(0) ~ "no disabilities",
    NIHSS_totalscore %in% c(1,2,3,4,5,6,7) ~ "one or more disabilities"
  ))

# T-test (non-parametric)
ttest_NIHSS <- wilcox.test(SF36_PF ~ NIHSS_grouped, data = df_stable_SF36)
print(ttest_NIHSS)

```

## Multivariate analysis
```{r}
set.seed(123)

# Multiple regression with VIF (Variance Inflation Factor); check collinearity 
# Regression model
model_stable <- lm(SF36_PF ~ age + sex + education_grouped + graad + KPS_grouped + NIHSS_grouped + RT_XT_treatment, data = df_stable_SF36)
summary(model_stable)

# VIF
vif_values_stable <- vif(model_stable)
print(vif_values_stable)

# Colinearity between histology & grade 
v_cramer <- assocstats(table(df_stable_SF36$histology, df_stable_SF36$graad))$cramer
table(v_cramer)

```





###### Part 3.3 ######

##### 3.3 CIS20 diagnosis; correlations (Pearson continue/independent sample t-test dichotomy/ANOVA categorical)

## Select data
```{r}
# Select CIS_fitness diagnosis
df_diagnosis_CIS_fitness <- df %>%
  filter(Time_moment == 1)
```

## Univariate analyses; check each variable separately 
# Age
```{r}
# Remove NA
df_clean_age <- na.omit(df_diagnosis_CIS_fitness[, c("age", "CIS_fitness")])
# Calculate Spearman-correlation between CIS_fitness and age, ignore NA
correlatie_CIS_fitness_age <- cor.test(df_clean_age$CIS_fitness, df_clean_age$age, method = "spearman", use = "complete.obs")
# Print results
print(paste("Spearman-correlatie between CIS_fitness and age:", correlatie_CIS_fitness_age))

rho <- correlatie_CIS_fitness_age$estimate
p_value <- correlatie_CIS_fitness_age$p.value

# Print results
print(paste("Spearman-correlatie tussen SF36_PF en leeftijd:", rho))
print(paste("P-waarde:", p_value))

```

# Sex
```{r}
# T-test (non-parametric)
ttest_sex <- wilcox.test(CIS_fitness ~ sex, data = df_diagnosis_CIS_fitness)
print(ttest_sex)

```

# Educational level
```{r}
# Convert to factor
df_diagnosis_CIS_fitness$education_old <- factor(df_diagnosis_CIS_fitness$education_old)

# Create new column with grouped values 
df_diagnosis_CIS_fitness <- df_diagnosis_CIS_fitness %>%
  mutate(education_grouped = case_when(
    education_old %in% c(1, 2) ~ "1-2",
    education_old %in% c(3, 4, 5) ~ "3-4-5",
    education_old %in% c(6, 7, 8) ~ "6-7-8"
  ))

# Convert to factor
df_diagnosis_CIS_fitness$education_grouped <- factor(df_diagnosis_CIS_fitness$education_grouped)

# ANOVA (non parametric)
anova_education <- kruskal.test(CIS_fitness ~ education_grouped, data = df_diagnosis_CIS_fitness)
print(anova_education)
summary(anova_education)
table(df_diagnosis_CIS_fitness$education_grouped)
# Post hoc
Posthoc_education <- dunn.test(df_diagnosis_CIS_fitness$CIS_fitness, df_diagnosis_CIS_fitness$education_grouped, method = "bonferroni")

```

# Histology
```{r}
# Convert to factor
df_diagnosis_CIS_fitness$histology <- factor(df_diagnosis_CIS_fitness$histology)

# ANOVA (non parametric)
anova_histology <- kruskal.test(CIS_fitness ~ histology, data = df_diagnosis_CIS_fitness)
print(anova_histology)
summary(anova_histology)
table(df_diagnosis_CIS_fitness$histology)
# Post hoc
Posthoc_histology <- dunn.test(df_diagnosis_CIS_fitness$CIS_fitness, df_diagnosis_CIS_fitness$histology, method = "bonferroni")

```

# Tumor location
```{r}
# Convert to factor
df_diagnosis_CIS_fitness$location <- factor(df_diagnosis_CIS_fitness$location)

# Create new column with grouped values 
df_diagnosis_CIS_fitness <- df_diagnosis_CIS_fitness %>%
  mutate(location_grouped = case_when(
    location %in% c(2) ~ "frontal",
    location %in% c(1, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 20, 21, 22, 23, 24, 25, 26) ~ "non-frontal"
  ))

# Convert to factor
df_diagnosis_CIS_fitness$location_grouped <- factor(df_diagnosis_CIS_fitness$location_grouped)

# T-test (non-parametric)
ttest_location <- wilcox.test(CIS_fitness ~ location_grouped, data = df_diagnosis_CIS_fitness)
print(ttest_location)

```

# Tumor lateralization (side)
```{r}
# Convert to factor
df_diagnosis_CIS_fitness$side <- factor(df_diagnosis_CIS_fitness$side)

# ANOVA
anova_side <- kruskal.test(CIS_fitness ~ side, data = df_diagnosis_CIS_fitness)
print(anova_side)
summary(anova_side)
table(df_diagnosis_CIS_fitness$side)
# Post hoc
Posthoc_side <- dunn.test(df_diagnosis_CIS_fitness$CIS_fitness, df_diagnosis_CIS_fitness$side, method = "bonferroni")

```

# WHO grade
```{r}
# Convert to factor
df_diagnosis_CIS_fitness$graad <- factor(df_diagnosis_CIS_fitness$graad)

# ANOVA
anova_grade <- kruskal.test(CIS_fitness ~ graad, data = df_diagnosis_CIS_fitness)
print(anova_grade)
summary(anova_grade)
table(df_diagnosis_CIS_fitness$graad)
# Post hoc
Posthoc_grade <- dunn.test(df_diagnosis_CIS_fitness$CIS_fitness, df_diagnosis_CIS_fitness$graad, method = "bonferroni")

```

# Type of surgery
```{r}
# T-test (non parametric)
ttest_surgery <- wilcox.test(CIS_fitness ~ OK_, data = df_diagnosis_CIS_fitness)
print(ttest_surgery)

```


# Presence of epilepsy
```{r}
# T-test (non parametric)
ttest_epilepsy <- wilcox.test(CIS_fitness ~ epilepsy_dich, data = df_diagnosis_CIS_fitness)
print(ttest_epilepsy)

```

# Use of dexamethasone
```{r}
# T-test (non parametric)
ttest_dexa <- wilcox.test(CIS_fitness ~ dexa, data = df_diagnosis_CIS_fitness)
print(ttest_dexa)

```

# KPS
```{r}
# KPS groups: = or < 70 and 80-90-100
df_diagnosis_CIS_fitness <- df_diagnosis_CIS_fitness %>%
  mutate(KPS_grouped = case_when(
    kps_total %in% c(80, 90, 100) ~ "80-90-100",
    kps_total %in% c(10, 20, 30, 40, 50, 60, 70) ~ "70 or lower"
  ))

# Convert to factor
df_diagnosis_CIS_fitness$KPS_grouped <- factor(df_diagnosis_CIS_fitness$KPS_grouped)

# T-test (non parametric)
ttest_KPS <- wilcox.test(CIS_fitness ~ KPS_grouped, data = df_diagnosis_CIS_fitness)
print(ttest_KPS)

```

# NIHSS score
```{r}
# NIHSS groups: = 0 or 1-6 
df_diagnosis_CIS_fitness <- df_diagnosis_CIS_fitness %>%
  mutate(NIHSS_grouped = case_when(
    NIHSS_totalscore %in% c(0) ~ "no disabilities",
    NIHSS_totalscore %in% c(1,2,3,4,5,6,7) ~ "one or more disabilities"
  ))

# Convert to factor
df_diagnosis_CIS_fitness$NIHSS_grouped <- factor(df_diagnosis_CIS_fitness$NIHSS_grouped)

# T-test (non-parametric)
ttest_NIHSS <- wilcox.test(CIS_fitness ~ NIHSS_grouped, data = df_diagnosis_CIS_fitness)
print(ttest_NIHSS)

```

## Multivariate analysis
```{r}
set.seed(123)

# Multiple regression with VIF (Variance Inflation Factor)
# Regression model
model_diagnosis <- lm(CIS_fitness ~ age + sex + education_grouped + side + KPS_grouped, data = df_diagnosis_CIS_fitness)
summary(model_diagnosis)

# VIF
vif_values_diagnosis <- vif(model_diagnosis)
print(vif_values_diagnosis)

```





###### Part 3.4 ######

##### 3.4 CIS20 stable disease; correlations (Pearson continue/independent sample t-test dichotomy/ANOVA categorical)
## Select data 
```{r}
# Select CIS_fitness stable phase 
df_stable_CIS_fitness <- df %>%
  filter(Time_moment == 2)
```

## Univariate analyses; check each variable separately 

# Age
```{r}
# Remove NA
df_clean_age <- na.omit(df_stable_CIS_fitness[, c("age", "CIS_fitness")])
# Calculate Spearman-correlation between CIS_fitness and age, ignore NA
correlatie_CIS_fitness_age <- cor.test(df_clean_age$CIS_fitness, df_clean_age$age, method = "spearman", use = "complete.obs")
# Print results
print(paste("Spearman-correlatie between CIS_fitness and age:", correlatie_CIS_fitness_age))

rho <- correlatie_CIS_fitness_age$estimate
p_value <- correlatie_CIS_fitness_age$p.value

# Print results
print(paste("Spearman-correlatie tussen SF36_PF en leeftijd:", rho))
print(paste("P-waarde:", p_value))

```

# Sex
```{r}
# T-test (non-parametric)
ttest_sex <- wilcox.test(CIS_fitness ~ sex, data = df_stable_CIS_fitness)
print(ttest_sex)

```

# Educational level
```{r}
# Convert to factor
df_stable_CIS_fitness$education_old <- factor(df_stable_CIS_fitness$education_old)

# Create new column with grouped values 
df_stable_CIS_fitness <- df_stable_CIS_fitness %>%
  mutate(education_grouped = case_when(
    education_old %in% c(1, 2) ~ "1-2",
    education_old %in% c(3, 4, 5) ~ "3-4-5",
    education_old %in% c(6, 7, 8) ~ "6-7-8"
  ))

# ANOVA (non parametric)
anova_education <- kruskal.test(CIS_fitness ~ education_grouped, data = df_stable_CIS_fitness)
print(anova_education)
summary(anova_education)
table(df_stable_CIS_fitness$education_grouped)
# Post hoc
Posthoc_education <- dunn.test(df_stable_CIS_fitness$CIS_fitness, df_stable_CIS_fitness$education_grouped, method = "bonferroni")

```

# Histology
```{r}
# Convert to factor
df_stable_CIS_fitness$histology <- factor(df_stable_CIS_fitness$histology)

# ANOVA (non parametric)
anova_location <- kruskal.test(CIS_fitness ~ histology, data = df_stable_CIS_fitness)
print(anova_location)
summary(anova_location)
table(df_stable_CIS_fitness$histology)
# Post hoc
Posthoc_location <- dunn.test(df_stable_CIS_fitness$CIS_fitness, df_stable_CIS_fitness$histology, method = "bonferroni")

```

# Tumor location
```{r}
# Convert to factor
df_stable_CIS_fitness$location <- factor(df_stable_CIS_fitness$location)

# Create new column with grouped values 
df_stable_CIS_fitness <- df_stable_CIS_fitness %>%
  mutate(location_grouped = case_when(
    location %in% c(2) ~ "frontal",
    location %in% c(1, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 20, 21, 22, 23, 24, 25, 26) ~ "non frontal"
  ))

# T-test (non-parametric)
ttest_location <- wilcox.test(CIS_fitness ~ location_grouped, data = df_stable_CIS_fitness)
print(ttest_location)

```

# Tumor lateralization (side)
```{r}
# Convert to factor
df_stable_CIS_fitness$side <- factor(df_stable_CIS_fitness$side)

# ANOVA
anova_side <- kruskal.test(CIS_fitness ~ side, data = df_stable_CIS_fitness)
print(anova_side)
summary(anova_side)
table(df_stable_CIS_fitness$side)
# Post hoc
Posthoc_side <- dunn.test(df_stable_CIS_fitness$CIS_fitness, df_stable_CIS_fitness$side, method = "bonferroni")

```

# WHO grade
```{r}
# Convert to factor
df_stable_CIS_fitness$graad <- factor(df_stable_CIS_fitness$graad)

# ANOVA
anova_grade <- kruskal.test(CIS_fitness ~ graad, data = df_stable_CIS_fitness)
print(anova_grade)
summary(anova_grade)
table(df_stable_CIS_fitness$graad)
# Post hoc
Posthoc_grade <- dunn.test(df_stable_CIS_fitness$CIS_fitness, df_stable_CIS_fitness$graad, method = "bonferroni")

```

# Type of surgery
```{r}
# T-test (non parametric)
ttest_surgery <- wilcox.test(CIS_fitness ~ OK_, data = df_stable_CIS_fitness)
print(ttest_surgery)

```

# Presence of epilepsy
```{r}
# T-test (non parametric)
ttest_epilepsy <- wilcox.test(CIS_fitness ~ epilepsy_dich, data = df_stable_CIS_fitness)
print(ttest_epilepsy)

```

# Use of dexamethasone
```{r}
# T-test (non parametric)
ttest_dexa <- wilcox.test(CIS_fitness ~ dexa, data = df_stable_CIS_fitness)
print(ttest_dexa)

```

# KPS
```{r}
# KPS groups: = or < 70 and 80-90-100
df_stable_CIS_fitness <- df_stable_CIS_fitness %>%
  mutate(KPS_grouped = case_when(
    kps_total %in% c(80, 90, 100) ~ "80-90-100",
    kps_total %in% c(10, 20, 30, 40, 50, 60, 70) ~ "70 or lower"
  ))

# Convert to factor
df_stable_CIS_fitness$KPS_grouped <- factor(df_stable_CIS_fitness$KPS_grouped)

# T-test (non parametric)
ttest_KPS <- wilcox.test(CIS_fitness ~ KPS_grouped, data = df_stable_CIS_fitness)
print(ttest_KPS)

```

# NIHSS score
```{r}
# NIHSS groups: = 0 or 1-6 
df_stable_CIS_fitness <- df_stable_CIS_fitness %>%
  mutate(NIHSS_grouped = case_when(
    NIHSS_totalscore %in% c(0) ~ "no disabilities",
    NIHSS_totalscore %in% c(1,2,3,4,5,6,7) ~ "one or more disabilities"
  ))

# Convert to factor
df_stable_CIS_fitness$NIHSS_grouped <- factor(df_stable_CIS_fitness$NIHSS_grouped)

# ANOVA
# T-test (non-parametric)
ttest_NIHSS <- wilcox.test(CIS_fitness ~ NIHSS_grouped, data = df_stable_CIS_fitness)
print(ttest_NIHSS)

```

# Radiotherapy & chemotherapy
```{r}
# Convert to factor
df_stable_CIS_fitness$RT_XT_treatment <- factor(df_stable_CIS_fitness$RT_XT_treatment)

# ANOVA
anova_RT_XT_treatment  <- kruskal.test(CIS_fitness ~ RT_XT_treatment , data = df_stable_CIS_fitness)
print(anova_RT_XT_treatment )
summary(anova_RT_XT_treatment )
table(df_stable_CIS_fitness$RT_XT_treatment )
# Post hoc
Posthoc_RT_XT_treatment  <- dunn.test(df_stable_CIS_fitness$CIS_fitness, df_stable_CIS_fitness$RT_XT_treatment , method = "bonferroni")

```

## Multivariate analysis
```{r}
set.seed(123)

# Multiple regression with VIF (Variance Inflation Factor); check collinearity 
# Regression model
model_stable <- lm(CIS_fitness ~ age + sex + education_grouped + KPS_grouped, data = df_stable_CIS_fitness)
summary(model_stable)

# VIF
vif_values_stable <- vif(model_stable)
print(vif_values_stable)

```










###### Part 3.5 ######

##### Part 3.5 SF36 difference diagnosis/stable; correlations

## Univariate analyses; linear mixed models; understanding how SF36 changes over time within individuals and how these changes vary with, for example, age. 
# Age
```{r}
# Fit linear mixed model; age
model_SF36_age <- lme(fixed = SF36_PF ~ Time_moment * age,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_SF36)

# Bekijk de model output
summary(model_SF36_age)

fixef(model_SF36_age)
model_summary_age <- summary(model_SF36_age)
model_summary_age$tTable

```

# Sex
```{r}
# Fit linear mixed model; sex
model_SF36_sex <- lme(fixed = SF36_PF ~ Time_moment * sex,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_SF36)

# Bekijk de model output
summary(model_SF36_sex)

fixef(model_SF36_sex)
model_summary_sex <- summary(model_SF36_sex)
model_summary_sex$tTable

```

# Education
```{r}
# Convert to factor
selected_pairs_SF36$education_grouped <- factor(selected_pairs_SF36$education_grouped)

# Fit linear mixed model; education
model_SF36_educ <- lme(fixed = SF36_PF ~ Time_moment * education_grouped,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_SF36)

# Bekijk de model output
summary(model_SF36_educ)

fixef(model_SF36_educ)
model_summary_educ <- summary(model_SF36_educ)
model_summary_educ$tTable

```

# Histology
```{r}
# Convert to factor
selected_pairs_SF36$histology <- factor(selected_pairs_SF36$histology)

# Fit linear mixed model; histology
model_SF36_hist <- lme(fixed = SF36_PF ~ Time_moment * histology,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_SF36)

# Bekijk de model output
summary(model_SF36_hist)

fixef(model_SF36_hist)
model_summary_hist <- summary(model_SF36_hist)
model_summary_hist$tTable

```

# Tumor location
```{r}
# Convert to factor
selected_pairs_SF36$location_grouped <- factor(selected_pairs_SF36$location_grouped)

# Fit linear mixed model; location
model_SF36_loc <- lme(fixed = SF36_PF ~ Time_moment * location_grouped,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_SF36)

# Bekijk de model output
summary(model_SF36_loc)

fixef(model_SF36_loc)
model_summary_loc <- summary(model_SF36_loc)
model_summary_loc$tTable

```

# Tumor lateralization
```{r}
# Convert to factor
selected_pairs_SF36$side <- factor(selected_pairs_SF36$side)

# Fit linear mixed model; side
model_SF36_side <- lme(fixed = SF36_PF ~ Time_moment * side,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_SF36)

# Bekijk de model output
summary(model_SF36_side)

fixef(model_SF36_side)
model_summary_side <- summary(model_SF36_side)
model_summary_side$tTable

```

# WHO grade
```{r}
# Convert to factor
selected_pairs_SF36$graad <- factor(selected_pairs_SF36$graad)

# Fit linear mixed model; WHO grade
model_SF36_graad <- lme(fixed = SF36_PF ~ Time_moment * graad,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_SF36)

# Bekijk de model output
summary(model_SF36_graad)

fixef(model_SF36_graad)
model_summary_graad <- summary(model_SF36_graad)
model_summary_graad$tTable

```

# Type of surgery
```{r}
## Type of surgery 
# only resection (no biopt)
```

# Epilepsy
```{r}
# Fit linear mixed model; epilepsy
model_SF36_epi <- lme(fixed = SF36_PF ~ Time_moment * epilepsy_dich,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_SF36)

# Bekijk de model output
summary(model_SF36_epi)

fixef(model_SF36_epi)
model_summary_epi <- summary(model_SF36_epi)
model_summary_epi$tTable

```

# Dexa
```{r}
# Fit linear mixed model; dexa
model_SF36_dexa <- lme(fixed = SF36_PF ~ Time_moment * dexa,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_SF36)

# Bekijk de model output
summary(model_SF36_dexa)

fixef(model_SF36_dexa)
model_summary_dexa <- summary(model_SF36_dexa)
model_summary_dexa$tTable

```

# Radiotherapy & chemotherapy
```{r}
# Convert to factor and set 'treatment 4 (nothing)' as reference level
selected_pairs_SF36$RT_XT_treatment <- factor(selected_pairs_SF36$RT_XT_treatment, levels = c("4", "1", "2", "3"))

# Fit linear mixed model; treatment
model_SF36_RTXT <- lme(fixed = SF36_PF ~ RT_XT_treatment * Time_moment,
                       random = ~ Time_moment | casenr, 
                       data = selected_pairs_SF36)

# Bekijk de model output
summary(model_SF36_RTXT)

fixef(model_SF36_RTXT)
model_summary_RTXT <- summary(model_SF36_RTXT)
model_summary_RTXT$tTable

```

# KPS
```{r}
# Convert to factor
selected_pairs_SF36$KPS_grouped <- factor(selected_pairs_SF36$KPS_grouped)

# Fit linear mixed model; KPS
model_SF36_kps <- lme(fixed = SF36_PF ~ Time_moment * KPS_grouped, 
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_SF36)

# Bekijk de model output
summary(model_SF36_kps)

fixef(model_SF36_kps)
model_summary_kps <- summary(model_SF36_kps)
model_summary_kps$tTable

```

# NIHSS score
```{r}
# Convert to factor
selected_pairs_SF36$NIHSS_grouped <- factor(selected_pairs_SF36$NIHSS_grouped)

# Fit linear mixed model; KPS
model_SF36_NIHSS <- lme(fixed = SF36_PF ~ Time_moment * NIHSS_grouped, 
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_SF36_NIHSS)

# Bekijk de model output
summary(model_SF36_NIHSS)

fixef(model_SF36_NIHSS)
model_summary_NIHSS <- summary(model_SF36_NIHSS)
model_summary_NIHSS$tTable

```

## Multivariate analysis; linear mixed model
```{r}
# Fit linear mixed model; multivariate; with significant outcomes & age, sex and education
model_SF36 <- lme(
  fixed = SF36_PF ~ Time_moment * (age + sex + education_grouped + RT_XT_treatment + graad + NIHSS_grouped),
  random = ~ Time_moment | casenr, 
  data = selected_pairs_SF36,
  na.action=na.exclude #NA wordt niet meegenomen in analyse (fitten van model), maar structuur blijft gelijk 
)

# Bekijk de model output
summary(model_SF36)

fixef(model_SF36)
model_summary <- summary(model_SF36)
model_summary$tTable

```





###### Part 3.6 ######

##### Part 3.6 CIS20 difference diagnosis/stable; correlations

## Univariate analyses; linear mixed models; understanding how CIS20 changes over time within individuals and how these changes vary with, for example, age. 
# Age
```{r}
# Fit linear mixed model; age
model_CIS_age <- lme(fixed = CIS_fitness ~ Time_moment * age,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_CIS)

# Bekijk de model output
summary(model_CIS_age)

fixef(model_CIS_age)
model_summary_age <- summary(model_CIS_age)
model_summary_age$tTable

```

# Sex
```{r}
# Fit linear mixed model; sex
model_CIS_sex <- lme(fixed = CIS_fitness ~ Time_moment * sex,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_CIS)

# Bekijk de model output
summary(model_CIS_sex)

fixef(model_CIS_sex)
model_summary_sex <- summary(model_CIS_sex)
model_summary_sex$tTable

```

# Education
```{r}
# Convert to factor
selected_pairs_CIS$education_grouped <- factor(selected_pairs_CIS$education_grouped)

# Fit linear mixed model; education
model_CIS_educ <- lme(fixed = CIS_fitness ~ Time_moment * education_grouped,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_CIS)

# Bekijk de model output
summary(model_CIS_educ)

fixef(model_CIS_educ)
model_summary_educ <- summary(model_CIS_educ)
model_summary_educ$tTable

```

# Histology
```{r}
# Convert to factor
selected_pairs_CIS$histology <- factor(selected_pairs_CIS$histology)

# Fit linear mixed model; histology
model_CIS_hist <- lme(fixed = CIS_fitness ~ Time_moment * histology,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_CIS)

# Bekijk de model output
summary(model_CIS_hist)

fixef(model_CIS_hist)
model_summary_hist <- summary(model_CIS_hist)
model_summary_hist$tTable

```

# Tumor location
```{r}
# Convert to factor
selected_pairs_CIS$location_grouped <- factor(selected_pairs_CIS$location_grouped)

# Fit linear mixed model; location
model_CIS_loc <- lme(fixed = CIS_fitness ~ Time_moment * location_grouped,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_CIS)

# Bekijk de model output
summary(model_CIS_loc)

fixef(model_CIS_loc)
model_summary_loc <- summary(model_CIS_loc)
model_summary_loc$tTable

```

# Tumor lateralization
```{r}
# Convert to factor
selected_pairs_CIS$side <- factor(selected_pairs_CIS$side)

# Fit linear mixed model; side
model_CIS_side <- lme(fixed = CIS_fitness ~ Time_moment * side,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_CIS)

# Bekijk de model output
summary(model_CIS_side)

fixef(model_CIS_side)
model_summary_side <- summary(model_CIS_side)
model_summary_side$tTable

```

# WHO grade
```{r}
# Convert to factor
selected_pairs_CIS$graad <- factor(selected_pairs_CIS$graad)

# Fit linear mixed model; WHO grade
model_CIS_graad <- lme(fixed = CIS_fitness ~ Time_moment * graad,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_CIS)

# Bekijk de model output
summary(model_CIS_graad)

fixef(model_CIS_graad)
model_summary_graad <- summary(model_CIS_graad)
model_summary_graad$tTable

```

# Type of surgery
```{r}
## Type of surgery 
# only resection (no biopt)
```

# Epilepsy
```{r}
# Fit linear mixed model; epilepsy
model_CIS_epi <- lme(fixed = CIS_fitness ~ Time_moment * epilepsy_dich,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_CIS)

# Bekijk de model output
summary(model_CIS_epi)

fixef(model_CIS_epi)
model_summary_epi <- summary(model_CIS_epi)
model_summary_epi$tTable

```

# Dexa
```{r}
# Fit linear mixed model; dexa
model_CIS_dexa <- lme(fixed = CIS_fitness ~ Time_moment * dexa,
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_CIS)

# Bekijk de model output
summary(model_CIS_dexa)

fixef(model_CIS_dexa)
model_summary_dexa <- summary(model_CIS_dexa)
model_summary_dexa$tTable

```

# Radiotherapy & chemotherapy
```{r}
# Convert to factor and set 'treatment 4 (nothing)' as reference level
selected_pairs_CIS$RT_XT_treatment <- factor(selected_pairs_CIS$RT_XT_treatment, levels = c("4", "1", "2", "3"))

# Fit linear mixed model; treatment
model_CIS20_RTXT <- lme(fixed = CIS_fitness ~ RT_XT_treatment * Time_moment,
                       random = ~ Time_moment | casenr, 
                       data = selected_pairs_CIS)

# Bekijk de model output
summary(model_CIS20_RTXT)

fixef(model_CIS20_RTXT)
model_summary_RTXT <- summary(model_CIS20_RTXT)
model_summary_RTXT$tTable

```

# KPS
```{r}
# Convert to factor
selected_pairs_CIS$KPS_grouped <- factor(selected_pairs_CIS$KPS_grouped)

# Fit linear mixed model; KPS
model_CIS_kps <- lme(fixed = CIS_fitness ~ Time_moment * KPS_grouped, 
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_CIS)

# Bekijk de model output
summary(model_CIS_kps)

fixef(model_CIS_kps)
model_summary_kps <- summary(model_CIS_kps)
model_summary_kps$tTable

```

# NIHSS score
```{r}
# Convert to factor
selected_pairs_CIS$NIHSS_grouped <- factor(selected_pairs_CIS$NIHSS_grouped)

# Fit linear mixed model; KPS
model_CIS_NIHSS <- lme(fixed = CIS_fitness ~ Time_moment * NIHSS_grouped, 
                           random = ~ Time_moment | casenr, 
                           data = selected_pairs_CIS_NIHSS)

# Bekijk de model output
summary(model_CIS_NIHSS)

fixef(model_CIS_NIHSS)
model_summary_NIHSS <- summary(model_CIS_NIHSS)
model_summary_NIHSS$tTable

```

## Multivariate analysis
```{r}
# Fit linear mixed model; multivariate; with significant outcomes & age, sex and education
model_CIS <- lme(
  fixed = CIS_fitness ~ Time_moment * (age + sex + education_grouped + RT_XT_treatment),
  random = ~ Time_moment | casenr, 
  data = selected_pairs_CIS
)

# Bekijk de model output
summary(model_CIS)

fixef(model_CIS)
model_summary <- summary(model_CIS)
model_summary$tTable

```




