---
title: "Towards unraveling fatigue"
author: "Jantine Röttgering"
date: "10-5-2022"
output: html_document
---

##### Email: j.rottgering@amsterdamumc.nl
##### Purpose of script: 
Analyses of the 2022 manuscript on fatigue in glioma patients.

##### Status: 
Under revision

##### Review History
Reviewed by: Vera Belgers
Date: 30-6-2022

##### Requirements
Created and ran with Rstudio 4.0.3, as a Markdown file.

#### Notes
This script is used for the analyses of the 2022 paper "Toward unraveling the association between fatigue and tumor- and treatment-related factors, symptoms, and quality of life in glioma". 

It consists of several parts:

1. Synthesizing mock data for the Glioma Cohort and the Longitudinal subgroup
2. Percentages of fatigued patients
3. Raincloudplots - Glioma Cohort
4. Linear mixed model - Glioma Cohort
5. Icc - Longitudinal subgroup
6. Change plot - Longitudinal subgroup
7. Multiple linear regression - Longitudinal subgroup

#### Load libraries

```{r, message=FALSE, warning = FALSE}
library(dplyr)
library(plyr)
library(ggplot2)
library(raincloudplots)
library(lme4)
library(irr)
library(car)
library(MASS)
library(MuMIn)
library(lmerTest)
```

# 1. Synthesizing mock datasets
The analyses in the 2022 fatigue paper consist of the Glioma Cohort and a subgroup, the Longitudinal subgroup. Here we provide mock data to run the analyses with. The data does not reflect the real distribution of the actual data, therefore the results produced in this script should merely be used to understand the analyses itself.

### Glioma cohort
```{r}
set.seed(1234)

df <- data.frame (random_id = c(1:222, 1:111),
                   assessment  = c(sample(c("Pre-operative", "Post-operative"), 333, replace=TRUE)),
                   CIS_total = c(sample(20:140, 333, replace=TRUE)),
                   CIS_fatigue  = c(sample(8:56, 333, replace=TRUE)),
                   CIS_concentration = c(sample(5:34, 333, replace=TRUE)),
                   CIS_motivation = c(sample(4:28, 333, replace=TRUE)),
                   CIS_activity = c(sample(3:21, 333, replace=TRUE)),
                   SF36_PF = c(sample(0:100, 333, replace=TRUE)),
                   SF36_SC = c(sample(0:100, 333, replace=TRUE)),
                   SF36_RLP = c(sample(0:100, 333, replace=TRUE)),
                   SF36_RLM = c(sample(0:100, 333, replace=TRUE)),
                   SF36_MH = c(sample(0:100, 333, replace=TRUE)),
                   SF36_P = c(sample(0:100, 333, replace=TRUE)),
                   SF36_HP = c(sample(0:100, 333, replace=TRUE)),
                   SF36_CIH = c(sample(0:100, 333, replace=TRUE)),
                   CESD_total = c(sample(0:100, 333, replace=TRUE)),
                   mos_cog = c(sample(11:36, 333, replace=TRUE)),
                   BN20_FU  = c(sample(0:100, 333, replace=TRUE)),
                   BN20_MD   = c(sample(0:100, 333, replace=TRUE)),
                   BN20_S  = c(sample(0:100, 333, replace=TRUE)),
                   side = c(sample(c("left", "right"), 333, replace=TRUE)),
                   kps = c(sample(c("70-80", "90-100"), 333, replace=TRUE)),
                   sex = c(sample(c("female", "male"), 333, replace=TRUE)),
                   age = c(sample(c(18:81), 333, replace=TRUE)),
                   verhage = c(sample(c("Low (1-4)", "Middle (5)", "High (6-7)"), 333, replace=TRUE)),
                   idh = c(sample(c("Mutated", "Not-mutated"), 333, replace=TRUE)),
                   location  = c(sample(c("Frontal", "Non-frontal"), 333, replace=TRUE)),
                   histology = c(sample(c("Astrocytoma ", "Oligodendroglioma", "Oligoastrocytoma", "Glioblastoma"), 333, replace=TRUE)),
                   dexa  = c(sample(c("No", "Yes"), 333, replace=TRUE)),
                   disease_status  = c(sample(c("Baseline", "Treatment", "Stable", "Progression"), 333, replace=TRUE)),
                   ct_before_assessment   = c(sample(c("No", "Yes"), 333, replace=TRUE)),
                   ct_before_assessment  = c(sample(c("No", "Yes"), 333, replace=TRUE)))
```

### Longitudinal subgroup
```{r}
set.seed(1234)

df_icc <- data.frame (random_id = c(1:70),
                   #preoperative PROMs  
                   CIS_total_preop = c(sample(20:140, 70, replace=TRUE)),
                   CIS_fatigue_preop  = c(sample(8:56, 70, replace=TRUE)),
                   CIS_concentration_preop = c(sample(5:34, 70, replace=TRUE)),
                   CIS_motivation_preop = c(sample(4:28, 70, replace=TRUE)),
                   CIS_activity_preop = c(sample(3:21, 70, replace=TRUE)),
                   SF36_PF_preop = c(sample(0:100, 70, replace=TRUE)),
                   SF36_SC_preop = c(sample(0:100, 70, replace=TRUE)),
                   SF36_RLP_preop = c(sample(0:100, 70, replace=TRUE)),
                   SF36_RLM_preop = c(sample(0:100, 70, replace=TRUE)),
                   SF36_MH_preop = c(sample(0:100, 70, replace=TRUE)),
                   SF36_P_preop = c(sample(0:100, 70, replace=TRUE)),
                   SF36_HP_preop = c(sample(0:100, 70, replace=TRUE)),
                   SF36_CIH_preop = c(sample(0:100, 70, replace=TRUE)),
                   CESD_total_preop = c(sample(0:100, 70, replace=TRUE)),
                   mos_cog_preop = c(sample(11:36, 70, replace=TRUE)),
                   BN20_FU_preop  = c(sample(0:100, 70, replace=TRUE)),
                   BN20_MD_preop   = c(sample(0:100, 70, replace=TRUE)),
                   BN20_S_preop  = c(sample(0:100, 70, replace=TRUE)),
                   #post-operative PROMs  
                   CIS_fatigue_postop  = c(sample(8:56, 70, replace=TRUE)),
                   #other variables
                   side = c(sample(c("left", "right"), 70, replace=TRUE)),
                   kps_preop = c(sample(c("70-80", "90-100"), 70, replace=TRUE)),
                   sex = c(sample(c("female", "male"), 70, replace=TRUE)),
                   age = c(sample(c(18:81), 70, replace=TRUE)),
                   verhage = c(sample(c("Low (1-4)", "Middle (5)", "High (6-7)"), 70, replace=TRUE)),
                   idh = c(sample(c("Mutated", "Not-mutated"), 70, replace=TRUE)),
                   location  = c(sample(c("Frontal", "Non-frontal"), 70, replace=TRUE)),
                   histology = c(sample(c("Astrocytoma ", "Oligodendroglioma", "Oligoastrocytoma", "Glioblastoma"), 70, replace=TRUE)),
                   dexa_preop  = c(sample(c("No", "Yes"), 70, replace=TRUE)),
                   ct_before_assessment_postop   = c(sample(c("No", "Yes"), 70, replace=TRUE)),
                   ct_before_assessment_postop  = c(sample(c("No", "Yes"), 70, replace=TRUE)))
```

# 2. Percentages of fatigued patients

Calculate the number of severely fatigued patients in the Glioma cohort
```{r}
#All patients
severely_fatigued_yes_no <- count(df$CIS_fatigue>34)
percentage_severly_fatigued<-(severely_fatigued_yes_no[2,2])/(severely_fatigued_yes_no[1,2]+severely_fatigued_yes_no[2,2])*100
percentage_severly_fatigued

#Preoperative
df_preok <- filter(df, assessment == "Pre-operative") 
severely_fatigued_yes_no_preok <- count(df_preok$CIS_fatigue>34)
percentage_severly_fatigued_preok<-(severely_fatigued_yes_no_preok[2,2])/(severely_fatigued_yes_no_preok[1,2]+severely_fatigued_yes_no_preok[2,2])*100
percentage_severly_fatigued_preok

#Postoperative
df_postok <- filter(df, assessment == "Post-operative")
severely_fatigued_yes_no_postok <- count(df_postok$CIS_fatigue>34)
percentage_severly_fatigued_postok<-(severely_fatigued_yes_no_postok[2,2])/(severely_fatigued_yes_no_postok[1,2]+severely_fatigued_yes_no_postok[2,2])*100
percentage_severly_fatigued_postok

# Treatment
treatment <- df %>% dplyr::filter(disease_status== "Treatment")
severely_fatigued_yes_no <- count(treatment$CIS_fatigue>34)
percentage_severly_fatigued_treatment<-(severely_fatigued_yes_no[2,2])/(severely_fatigued_yes_no[1,2]+severely_fatigued_yes_no[2,2])*100
percentage_severly_fatigued_treatment

# Stable disease
stable <- df %>% dplyr::filter(disease_status== "Stable")
severely_fatigued_yes_no <- count(stable$CIS_fatigue>34)
percentage_severly_fatigued_stable<-(severely_fatigued_yes_no[2,2])/(severely_fatigued_yes_no[1,2]+severely_fatigued_yes_no[2,2])*100
percentage_severly_fatigued_stable

# Progression
progression <- df %>% dplyr::filter(disease_status== "Progression")
severely_fatigued_yes_no <- count(progression$CIS_fatigue>34)
percentage_severly_fatigued_progression <-(severely_fatigued_yes_no[2,2])/(severely_fatigued_yes_no[1,2]+severely_fatigued_yes_no[2,2])*100
percentage_severly_fatigued_progression
```

Calculate the number of severely fatigued patients in the Longitudinal subgroup
```{r}
#Preoperative
severely_fatigued_yes_no_preok <- count(df_icc$CIS_fatigue_preop>34)
percentage_severly_fatigued_preok<-(severely_fatigued_yes_no_preok[2,2])/(severely_fatigued_yes_no_preok[1,2]+severely_fatigued_yes_no_preok[2,2])*100
percentage_severly_fatigued_preok

#Postoperative
severely_fatigued_yes_no_postok <- count(df_icc$CIS_fatigue_postop>34)
percentage_severly_fatigued_postok<-(severely_fatigued_yes_no_postok[2,2])/(severely_fatigued_yes_no_postok[1,2]+severely_fatigued_yes_no_postok[2,2])*100
percentage_severly_fatigued_postok
```

# 3. Raincloudplots Glioma Cohort

Used repository: https://github.com/jorvlan/raincloudplots

Here the code for a raincloudplot is provided. Change the variable according to the desired variable.
In the paper we present the variabels: CIS_total, CIS_fatigue, CIS_motivation, CIS_concentration, CIS_activity.

#### Format data
```{r}
df_1x1 <- data_1x1(
  array_1 = df_preok$CIS_total, #change variable
  array_2 = df_postok$CIS_total, #change variable
  jit_distance = .09,
  jit_seed = 321)
```

#### Make the figure
```{r}
raincloud_figure <- raincloud_1x1_repmes(
  data = df_1x1,
  colors = (c('#73716c','#73716c')), 
  fills = (c('#73716c','#73716c')), 
  line_color = '#73716c',
  line_alpha = .4,
  size = 2,
  alpha = .4,
  align_clouds = FALSE) +
  geom_hline(yintercept=76,linetype="dashed", color="#73716c")+  
  scale_x_continuous(breaks=c(0.7,2.3), labels=c("Preoperative", "Postoperative"), limits=c(0, 3)) + 
  scale_y_continuous(limits=c(20, 140)) + #change scale y-axis
  xlab(NULL)+
  ylab("CIS total score") + #change ylab
  theme_classic(base_size = 18)

raincloud_figure
```

# 4. Linear mixed model Glioma Cohort

Check what probability distribution best fits the CIS_fatigue: 

```{r}
df$CIS_fatigue.t <- df$CIS_fatigue + 1
qqp(df$CIS_fatigue.t, "norm")
qqp(df$CIS_fatigue.t, "lnorm")

nbinom <- fitdistr(df$CIS_fatigue.t, "Negative Binomial")
qqp(df$CIS_fatigue.t, "nbinom", size = nbinom$estimate[[1]], mu = nbinom$estimate[[2]])

poisson <- fitdistr(df$CIS_fatigue.t, "Poisson")
qqp(df$CIS_fatigue.t, "pois", poisson$estimate)

gamma <- fitdistr(df$CIS_fatigue.t, "gamma")
qqp(df$CIS_fatigue.t, "gamma", shape = gamma$estimate[[1]], rate = gamma$estimate[[2]])
```
Indeed a normal distribution

#### STEP 1 LMM: for each individual variable, check if there is an association. (See table 2 in manuscript)

If there is an association, then this variable is used in step 2.
```{r, message=FALSE}
vars = c("SF36_PF", "SF36_SC", "SF36_RLP", "SF36_RLM", "SF36_MH", "SF36_P", "SF36_HP", "SF36_CIH", "CESD_total", "mos_cog",  "BN20_FU", "BN20_MD", "BN20_S", "side", "kps", "sex", "age", "verhage", "idh", "location", "assessment", "histology", "dexa" , "disease_status", "ct_before_assessment", "ct_before_assessment")

#lapply loop to compute a linear mixed model (lmm) for all the variables separately
lmm_individual_variable = lapply(setNames(vars, vars), function(var) {
  form = paste("CIS_fatigue ~ ", var, "+ (1|random_id)")
  lmer(form, data=df,REML = FALSE)
})

summary(lmm_individual_variable[[2]])  #change the number to see the other variabels. Number of myvars=26.
```

See table 2 in manuscript. Variables to exclude for the next step: age, dexa, location, idh, histology, disease_status, ct_before_assessment, ct_before_assessment, grade assessment. Of course, since this script runs with mock data, the p-values and estimates so not reflect the values in the manuscript. 

#### STEP 2 LMM
This model includes the variables that show a significant univariate association. 
```{r}
no.na.data <- na.omit(df[c("CIS_fatigue", "sex", "CESD_total",  "mos_cog","kps", "SF36_PF","SF36_MH" ,"SF36_HP", "SF36_SC","SF36_P","SF36_RLP","SF36_RLM" ,"SF36_CIH","BN20_MD","BN20_S" ,"BN20_FU", "verhage","side", "random_id")])

modelall <- lmer(CIS_fatigue ~ 
sex+ CESD_total+mos_cog+ kps+ SF36_PF+ SF36_MH + SF36_HP+ SF36_SC+ SF36_P +  
SF36_RLP+ SF36_RLM +SF36_CIH+BN20_MD +BN20_S +BN20_FU+ verhage+side
                   + (1 | random_id), data = no.na.data,
    REML = FALSE)

summary(modelall)
```

#### STEP 3: the final model - LMM

Perform backward selection using the step test from the lmerTest package.
```{r}
step_model <- step(modelall)
step_model
```

```{r}
finalmodel <- lmer(CIS_fatigue ~ 
SF36_RLP+ 
  SF36_PF+
  SF36_HP+
  SF36_SC+
  CESD_total+ 
  side
                 + (1 | random_id), data = df,
    REML = FALSE)

summary(finalmodel)
r.squaredGLMM(finalmodel)
```

See Table 3. Determinants of fatigue severity in manuscript.

# 5. ICC longitudinal subgroup 

```{r}
icc(df_icc[ , c("CIS_fatigue_preop", "CIS_fatigue_postop") ], model="twoway", type="agreement", unit="single")
```

# 6. Change plot longitudinal subgroup 
Regarding the Longitudinal subgroup.

##### Make a delta score and define whether values can be classified as change or not
```{r}
df_icc$delta_cisf = df_icc$CIS_fatigue_preop - df_icc$CIS_fatigue_postop

df_icc$Legend = ifelse(df_icc$delta_cisf >6.9, "Clinically significant improvement",
                  ifelse(df_icc$delta_cisf <(-6.9), "Deteroration", 
                  ifelse(df_icc$delta_cisf<7 & df_icc$delta_cisf> (-7), "Minimal change", "Unknown")))
count(df_icc$Legend)
```

##### Fig 2. Change in fatigue severity scores between baseline and second assessment
```{r}
df_plot<- df_icc %>% dplyr::select(CIS_fatigue_preop, CIS_fatigue_postop, Legend) %>% na.omit()

change_plot <- ggplot(df_plot, aes(CIS_fatigue_preop, CIS_fatigue_postop))+
    geom_point(aes(colour = Legend, shape = Legend), position = "jitter", size = 3) +
    geom_hline(yintercept=35,linetype="dashed", color="darkgrey")+  
    geom_vline(xintercept=35,linetype="dashed", color="darkgrey")+  
  xlab("Pre-surgery CIS fatigue severity")+
  ylab("Post-surgery CIS fatigue assessment")+
  theme_classic()

change_plot + scale_color_manual(values=c("#3c3b3d","#000000","#b8b6ba"))
```


# 7. Multiple linear regression

Find predictors of postoperative CIS-fatigue with the df_icc = longitudinal subgroup.

#### STEP 1 MLR
```{r}
myvars <- c("CIS_fatigue_preop", "CIS_concentration_preop", "CIS_motivation_preop", "CIS_activity_preop", "SF36_PF_preop", "SF36_SC_preop", "SF36_RLP_preop", "SF36_RLM_preop", "SF36_MH_preop", "SF36_P_preop", "SF36_HP_preop", "SF36_CIH_preop", "mos_cog_preop", "CESD_total_preop", "BN20_FU_preop", "BN20_MD_preop", "BN20_S_preop", "sex", "age", "verhage", "idh", "side","location", "histology", "dexa_preop" , "kps_preop", "ct_before_assessment_postop", "ct_before_assessment_postop")

linear_regression_individual_variable <- lapply(myvars, function(dvar) lm(eval(paste0(' CIS_fatigue_postop ~ ', dvar)), data = df_icc))

summary(linear_regression_individual_variable[[1]]) #change the number to see the other variabels. Number of myvars=28.
```

Variables to exclude for the next step: BN20_S_preop, age, verhage, sex, idh, side, location, histology, dexa_preop, kps_preop, ct_before_assessment_postop, radiotherapy

#### STEP 2 MLR
Before backward selection:
```{r}
no.na.data_icc <- na.omit(df_icc[c("CIS_fatigue_postop", "CIS_fatigue_preop", "CIS_concentration_preop", "CIS_motivation_preop", "CIS_activity_preop", "SF36_PF_preop", "SF36_SC_preop", "SF36_RLP_preop","SF36_RLM_preop", "SF36_MH_preop", "SF36_P_preop", "SF36_HP_preop", "SF36_CIH_preop", "mos_cog_preop", "CESD_total_preop", "BN20_FU_preop", "BN20_MD_preop")])

model <- lm(CIS_fatigue_postop ~
CIS_fatigue_preop+
CIS_concentration_preop+
CIS_motivation_preop+
CIS_activity_preop+
SF36_PF_preop+
SF36_SC_preop+  
SF36_RLP_preop+
SF36_RLM_preop+
SF36_MH_preop+
SF36_P_preop+
SF36_HP_preop+
SF36_CIH_preop+  
mos_cog_preop+
CESD_total_preop+
BN20_FU_preop+
BN20_MD_preop,
            data = no.na.data_icc) 

summary(model)
```

# STEP 3: backward selection

```{r}
backward_lmr_model <- stepAIC(model, direction = "backward", 
                      trace = FALSE)

backward_lmr_model$coefficients #variables to include in final model
```

The final model after backward selection:
```{r}
final_model <- lm(CIS_fatigue_postop ~
CIS_fatigue_preop+
SF36_PF_preop,
            data = df_icc) 

summary(final_model)
```
